DATA 621 - Business Analytics and Data Mining

### 2020-11-10 HW4/HW4_Draft.Rmd
We don't need the entire tidyverse, just dplyr. That will speed things up. Must load dplyr after arm, otherwise select from MASS masks select from dplyr. Oh, and copyedits.

```{r loadData, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(mice)
library(arm)
library(corrplot)
library(data.table)

@@

One thing to note about Model 3 is that it does *not* assume that an `NA` in 
the `HOME_VAL` variable is indicative of renting. This is a key difference between
this model and the previous 2.

```{r m3clean}
# Model 3 - data transformation
# MICE imputation on train and test set
m3train <- complete(mice(data = trnX %>% select(-INDEX), method = "pmm", print=FALSE),3)
m3train$TARGET_FLAG <- factor(m3train$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))

m3test <- complete(mice(data = tstX %>% select(-INDEX), method = "pmm", print=FALSE),3)
m3test$TARGET_FLAG <- factor(m3test$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))

```

# BUILD MODELS
## Model Set 1
### Dummy Variables  
As there are many factor variables, dummy variables will be created. The target
variables will be removed and held to the side so as not to contaminate the
fitting.

@@

Unlike the frequency model, many of the resulting predictors are not significant
at even the 10% level. What this may be telling us is that linear regression is
***not*** the proper model for this analysis and the algorithm is struggling to
add non-linearity via the interactions. In practice, at the very least a GLM
with a gamma link would be tried or, more sophisticated predictive models such
as tree-based ones would be investigated. Most likely, MLE would be used to
estimate the severity and a hierarchial contingent stochastic cash flow model
would be built such that given a claim, a loss would be generated from a much
more amenable distribution, such as a Pareto or a Burr.

@@

```{r model3RunBaseLog}
# Base log model 3
m3BaseLog <- glm(TARGET_FLAG~., family = binomial, data = m3train %>% select(-TARGET_AMT))
kable(summary(m3BaseLog)$coefficients, digits = 3L,
      caption = 'Model 3 - Base Logistic Regression Output')

```

@@

```{r model3RunFreq}
# Final log model 3
train_control <- trainControl(method = "cv", number = 10)
m3FreqFit <- train(TARGET_FLAG ~ KIDSDRIV+INCOME+PARENT1+HOME_VAL+MSTATUS+SEX+EDUCATION+JOB+
              TRAVTIME+CAR_USE+BLUEBOOK+TIF+CAR_TYPE+OLDCLAIM+CLM_FREQ+REVOKED+MVR_PTS+
              URBANICITY,
              data = m3train %>% select(-TARGET_AMT),
              trControl = train_control,
              method = "glm",
              family=binomial())
kable(summary(m3FreqFit)$coefficients, digits = 3L,
      caption = 'Model 3 - Final Logistic Regression Output')

```

@@

#### Base Model
We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset, but only for the records that have
a TARGET_AMT > $0.

@@

m3BaseLin <- lm(TARGET_AMT ~ ., data=m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))
knitr::kable(summary(m3BaseLin)$coefficients, digits = 3L,
             caption = 'Model 3 - Base Linear Regression Output')

```

@@

m3SevFit <- lm(TARGET_AMT~ BLUEBOOK+CAR_AGE+CAR_TYPE, data = m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))

knitr::kable(summary(m3SevFit)$coefficients, digits = 3L,
             caption = 'Model 3 - Final Linear Regression Output')

```

@@

* The positive coefficient for the `BLUEBOOK` variable means that the higher
the bluebook value for the vehicle, the more the payout will be. This makes
sense. 
* The negative coefficient for the `CAR_AGE` variable means that the older
the vehicle is, the less the payout is. Once again, this makes sense. 
* Non-intuitively, the positive coefficient on the Van option for the 
`CAR_TYPE` variable means that van collisions have a higher payout. All 
other options for this variable (including a sportscar) result in a 
lower payout.

>>>>

```{r loadData, include=FALSE}
# Load necessary libraries
library(arm)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(mice)
library(corrplot)
library(data.table)

@@

One thing to note about Model 3 is that it does *not* assume that an `NA` in the
`HOME_VAL` variable is indicative of renting. This is a key difference between
this model and the previous two models.

```{r m3clean}
# Model 3 - data transformation
# MICE imputation on train and test set
m3train <- complete(mice(data = trnX %>% select(-INDEX),
                         method = "pmm", print = FALSE), 3)
m3train$TARGET_FLAG <- factor(m3train$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))

m3test <- complete(mice(data = tstX %>% select(-INDEX),
                        method = "pmm", print = FALSE), 3)
m3test$TARGET_FLAG <- factor(m3test$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))
```

# BUILD MODELS
## Model Set #1
### Dummy Variables  
As there are many factor variables, dummy variables will be created. The target
variables will be removed and held to the side so as not to contaminate the
fitting.

@@

Unlike the frequency model, many of the resulting predictors are not significant
at even the 10% level. What this may be telling us is that linear regression is
***not*** the proper model for this analysis and the algorithm is struggling to
add non-linearity via the interactions. In practice, at the very least a GLM
with a gamma link would be tried or, more sophisticated predictive models such
as tree-based ones would be investigated. Most likely, MLE would be used to
estimate the severity and a hierarchical contingent stochastic cash flow model
would be built such that given a claim, a loss would be generated from a much
more amenable distribution, such as a Pareto or a Burr.

@@

```{r model3RunBaseLog}
# Base log model 3
m3BaseLog <- glm(TARGET_FLAG ~ ., family = binomial,
                 data = m3train %>% select(-TARGET_AMT))
kable(summary(m3BaseLog)$coefficients, digits = 3L,
      caption = 'Model 3 - Base Logistic Regression Output')
```

@@

```{r model3RunFreq}
# Final log model 3
train_control <- trainControl(method = "cv", number = 10)
m3FreqFit <- train(TARGET_FLAG ~ KIDSDRIV + INCOME + PARENT1 + HOME_VAL +
                     MSTATUS + SEX + EDUCATION + JOB + TRAVTIME + CAR_USE +
                     BLUEBOOK + TIF + CAR_TYPE + OLDCLAIM + CLM_FREQ + REVOKED +
                     MVR_PTS + URBANICITY,
              data = m3train %>% select(-TARGET_AMT),
              trControl = train_control,
              method = "glm",
              family=binomial())
kable(summary(m3FreqFit)$coefficients, digits = 3L,
      caption = 'Model 3 - Final Logistic Regression Output')
```

@@

#### Base Model
We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset, but only for the records that have
a `TARGET_AMT` > $0.

@@

m3BaseLin <- lm(TARGET_AMT ~ .,
                data = m3train %>%
                  filter(TARGET_AMT > 0) %>%
                  select(-TARGET_FLAG))
knitr::kable(summary(m3BaseLin)$coefficients, digits = 3L,
             caption = 'Model 3 - Base Linear Regression Output')
```

@@

m3SevFit <- lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + CAR_TYPE,
               data = m3train %>%
                 filter(TARGET_AMT > 0) %>%
                 select(-TARGET_FLAG))

knitr::kable(summary(m3SevFit)$coefficients, digits = 3L,
             caption = 'Model 3 - Final Linear Regression Output')
```

@@

  * The positive coefficient for the `BLUEBOOK` variable means that the higher
  the bluebook value for the vehicle, the more the payout will be. This makes
  sense. 
  * The negative coefficient for the `CAR_AGE` variable means that the older the
  vehicle is, the less the payout is. Once again, this makes sense. 
  * Non-intuitively, the positive coefficient on the Van option for the
  `CAR_TYPE` variable means that van collisions have a higher payout. All other
  options for this variable (including a sports car) result in a lower payout.
